---
title: "Getting Started with pgeeVar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with pgeeVar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows a basic `pgeeVar` analysis on one simulated binary
longitudinal dataset. It is a small working example, not a reproduction of the
manuscript analyses.

The package workflow is centered on three tasks:

1. simulate correlated binary longitudinal data;
2. fit a penalized GEE model;
3. extract one or more covariance estimators.

## Simulate a working dataset

`simulate_correlated_binary_data()` is the user-facing data generator. It
returns an analysis-ready data frame and stores the generating setup in a
`"pgee_scenario"` attribute.

```{r}
library(pgeeVar)

example_data <- simulate_correlated_binary_data(
  N = 20,
  gamma = 0.5,
  n_i = 4,
  beta = c(-1.2, 0.6, 0.3),
  rho = 0.2,
  corstr = "exchangeable",
  seed = 1
)

simulation_metadata <- attr(example_data, "pgee_scenario")

head(example_data)
simulation_metadata[c("N", "rho", "corstr", "clf_ok", "event_rate")]
range(simulation_metadata$n_i)
```

The simulated data frame contains the cluster identifier `id`, the binary
response `y`, and the covariates used in the fitted model. The scenario
attribute is useful in simulation work because it lets you keep the generating
setup attached to the data without managing a separate object.

## Fit a penalized GEE model

`pgee()` is the main modeling interface. It follows an `lm()` or `glm()`
pattern: formula, data frame, and cluster identifier.

```{r}
fitted_model <- pgee(
  y ~ X1 + obstime,
  data = example_data,
  id = id,
  corstr = "exchangeable"
)

fitted_model
model_summary <- summary(fitted_model)
model_summary
```

The printed fit is short. The summary adds the coefficient table, working
correlation estimate, convergence information, and the default covariance
choice.

If you want the coefficient table as a regular data frame for a report or a
small diagnostic check, extract it directly from the summary object.

```{r}
coefficient_table <- data.frame(
  term = rownames(model_summary$coefficients),
  model_summary$coefficients,
  row.names = NULL
)

coefficient_table
```

## Inspect the available covariance estimators

The package currently exposes the 14 estimators studied in the manuscript.
The public package interface uses short estimator names and keeps the
manuscript notation in a separate catalog column.

```{r}
estimator_catalog <- available_estimators()
estimator_catalog[, c("name", "notation", "class", "requires_balance")]
```

If you only need the accepted estimator names, use
`available_estimator_names()`.

```{r}
available_estimator_names()
```

## Compare a few covariance estimators

`vcov()` uses the fit's default covariance estimator unless you request another
type. `compute_pgee_variances()` is the lower-level helper that returns the
computed covariance matrices directly.

```{r}
ar_covariance <- vcov(fitted_model, type = "AR")
mbn_covariance <- vcov(fitted_model, type = "MBN")
lz_covariance <- vcov(fitted_model, type = "LZ")

se_table <- data.frame(
  term = names(coef(fitted_model)),
  LZ = sqrt(diag(lz_covariance)),
  MBN = sqrt(diag(mbn_covariance)),
  AR = sqrt(diag(ar_covariance)),
  row.names = NULL
)

se_table
```

That table is often enough for a first comparison. If you want a selected subset
of covariance matrices for later use, request them directly.

```{r}
selected_covariances <- compute_pgee_variances(fitted_model, c("MD", "FZ", "AR"))
names(selected_covariances)
dim(selected_covariances$AR)
```

## Predictions, residuals, and intervals

The `"pgee"` object supports the standard extraction methods needed for routine
analysis.

```{r}
predict(fitted_model, type = "response")[1:6]
residuals(fitted_model, type = "pearson")[1:6]
confint(fitted_model, type = "AR")
```

Predictions on new data use the stored formula terms and factor levels, so you
can pass a subset of the original data frame or a new data frame with the same
variables.

```{r}
prediction_rows <- example_data[1:4, ]
predict(fitted_model, newdata = prediction_rows, type = "response")
```

## When to drop to the low-level engine

Most users should stay with `simulate_correlated_binary_data()` and `pgee()`.
The lower-level functions `dgp_generate()` and `pgee_fit()` are still available
for simulation scripts that need tighter control over the generated object or
over the fitting inputs.

```{r}
set.seed(2)

generated_data <- dgp_generate(
  N = 20,
  gamma = 0.5,
  n_i = 4,
  beta = c(-1.2, 0.6, 0.3),
  rho = 0.2
)

low_level_fit <- pgee_fit(
  y = generated_data$data$y,
  x = generated_data$data[generated_data$covnames],
  id = generated_data$data$id,
  corr_type = generated_data$corr_type
)

names(compute_pgee_variances(low_level_fit, c("LZ", "AR")))
```

Use the low-level path when you need raw control over the generated object or
want to trim wrapper overhead inside a larger simulation loop. For ordinary data
analysis, `simulate_correlated_binary_data()` and `pgee()` are the simpler path.
