---
title: "Simulation Design Workflow with pgeeVar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation Design Workflow with pgeeVar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to set up a small simulation loop with `pgeeVar`. The
examples are intentionally short so the vignette can build quickly. Full
manuscript-scale simulations belong in repository scripts, not in the package
vignettes.

## A small simulation helper

The helper below simulates one dataset, fits a penalized GEE model, and returns
the `X1` estimate together with two covariance-based standard errors. That is
enough structure for a compact design check or a unit-level simulation summary.

```{r}
library(pgeeVar)

run_single_replication <- function(seed, N, corstr, rho, n_i = 4) {
  simulated_data <- simulate_correlated_binary_data(
    N = N,
    gamma = 0.5,
    n_i = n_i,
    beta = c(-1.2, 0.6, 0.3),
    rho = rho,
    corstr = corstr,
    seed = seed
  )

  fitted_model <- pgee(
    y ~ X1 + obstime,
    data = simulated_data,
    id = id,
    corstr = corstr
  )

  ar_standard_error <- sqrt(diag(vcov(fitted_model, type = "AR")))["X1"]
  mbn_standard_error <- sqrt(diag(vcov(fitted_model, type = "MBN")))["X1"]

  c(
    x1_estimate = unname(coef(fitted_model)["X1"]),
    ar_standard_error = unname(ar_standard_error),
    mbn_standard_error = unname(mbn_standard_error),
    converged = as.numeric(fitted_model$converged)
  )
}
```

## Compare a few small scenarios

The next chunk runs a compact design with ten seeds per scenario. That is
enough to show the workflow and still keep the vignette light.

```{r}
simulation_conditions <- list(
  exchangeable = list(N = 20, corstr = "exchangeable", rho = 0.2),
  ar1 = list(N = 20, corstr = "ar1", rho = 0.2)
)

replication_ids <- 1:10

summarize_condition <- function(spec) {
  condition_results <- do.call(
    rbind,
    lapply(replication_ids, function(seed) {
      run_single_replication(
        seed = seed,
        N = spec$N,
        corstr = spec$corstr,
        rho = spec$rho
      )
    })
  )

  data.frame(
    mean_x1_estimate = mean(condition_results[, "x1_estimate"]),
    sd_x1_estimate = stats::sd(condition_results[, "x1_estimate"]),
    mean_ar_standard_error = mean(condition_results[, "ar_standard_error"]),
    mean_mbn_standard_error = mean(condition_results[, "mbn_standard_error"]),
    convergence_rate = mean(condition_results[, "converged"])
  )
}

condition_summary <- do.call(
  rbind,
  lapply(simulation_conditions, summarize_condition)
)

condition_summary
```

Each row is a simple condition-level summary. In a larger study you would
usually add bias, empirical standard deviation, and coverage summaries, but the
structure is the same.

This style of helper is enough for many package-level simulation checks:

- keep the data generation in one place;
- keep the model formula fixed within a block of scenarios;
- store only the summaries you need for the design question at hand.

## Working with unbalanced designs

Some estimators in `pgeeVar` require balanced cluster sizes. On an unbalanced
dataset, those estimators still have named slots in the output, but the
returned matrices are filled with `NA`. The first call below shows the
documented warning once.

```{r}
unbalanced_example_data <- simulate_correlated_binary_data(
  N = 18,
  gamma = 0.5,
  n_i = c(3, 5),
  beta = c(-1.2, 0.6, 0.3),
  rho = 0.2,
  corstr = "exchangeable",
  seed = 11
)

unbalanced_fit <- pgee(
  y ~ X1 + obstime,
  data = unbalanced_example_data,
  id = id,
  corstr = "exchangeable"
)

pan_covariance <- vcov(unbalanced_fit, type = "Pan")
gst_covariance <- suppressWarnings(vcov(unbalanced_fit, type = "GST"))

all(is.na(pan_covariance))
all(is.na(gst_covariance))
```

You can use `available_estimators()` to see which estimators require balance
before you launch a larger simulation grid.

```{r}
subset(available_estimators(), requires_balance)
```

## Using the low-level engine in a script

For larger simulation studies, you may prefer the lower-level engine because it
returns the full generating object and the raw fitter output.

```{r}
set.seed(3)

generated_data <- dgp_generate(
  N = 20,
  gamma = 0.5,
  n_i = 4,
  beta = c(-1.2, 0.6, 0.3),
  rho = 0.2,
  corr_type = "exchangeable"
)

low_level_fit <- pgee_fit(
  y = generated_data$data$y,
  x = generated_data$data[generated_data$covnames],
  id = generated_data$data$id,
  corr_type = generated_data$corr_type
)

selected_covariances <- compute_pgee_variances(low_level_fit, c("LZ", "MBN", "AR"))
names(selected_covariances)
```

The wrapper path and the low-level path can coexist in the same project. A
common pattern is to prototype with `pgee()`, then switch the inner loop to
`pgee_fit()` once you need tighter control over the stored fit object.

## Practical limits for package vignettes

The examples here are intentionally small. For manuscript work:

- use more seeds than the vignette uses here;
- separate balanced and unbalanced designs clearly;
- in manuscript-scale runs, log both convergence and warning behavior, not only
  coefficient summaries;
- keep the package functions as the canonical implementation once the package
  surface is frozen.
